Taking Linear Logic Apart

### 2018-05-01 doc/tlla2018/main.tex
minor fix
commented out conclusions

\paragraph{Types}
Channels in \cp are typed using a session type system which corresponds to classical linear logic. A process communicating over a channel of type \ty{A \tens B} must be ready to split into two independent processes (two processes which share no channels), ready to communicate over a channel of type \ty{A} and \ty{B}, respectively, and process communicating over a channel of type \ty{A \parr B} is dual to this, and therefore must be ready to start communicating concurrently with two independent processes. A process communicating over a channel of type \ty{A \plus B} can choose to communicate either as \ty{A} or \ty{B}, and a process communicating over a channel of type \ty{A \with B} must be able to communicate either as \ty{A} or \ty{B}.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}\noindent
Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
Environments associate channels with types. Names in environments must be unique, and environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$. We write $\ty{\cdot}$ for the empty environment.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}
  \)
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\cpInfAx \cpInfCut \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfTens \cpInfParr \cpInfOne \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfBot \cpInfPlus1 \cpInfPlus2\end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfWith \cpInfNil \cpInfTop \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}

Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}\noindent

@@

\section{Conclusions}
\label{sec:conclusions}

We introduced Hypersequent Classical Processes, a process calculus which addresses the mismatch between proof terms and \textpi-calculus terms which exists in Classical Processes~\citep{wadler2012}. We showed that \hcp enjoys preservation, progress and termination, and described its relation to \cp.

>>>>

\paragraph{Types}
Channels in \cp are typed using a session type system which corresponds to classical linear logic. A process communicating over a channel of type \ty{A \tens B} must be ready to split into two independent processes (two processes which share no channels), ready to communicate over a channel of type \ty{A} and \ty{B}, respectively, and process communicating over a channel of type \ty{A \parr B} is dual to this, and therefore must be ready to start communicating concurrently with two independent processes. A process communicating over a channel of type \ty{A \plus B} can choose to communicate either as \ty{A} or \ty{B}, and a process communicating over a channel of type \ty{A \with B} must be able to communicate either as \ty{A} or \ty{B}.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}\noindent
Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
Environments associate channels with types. Names in environments must be unique, and environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$. We write $\ty{\cdot}$ for the empty environment.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}
  \)
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\cpInfAx \cpInfCut \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfTens \cpInfParr \cpInfOne \end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfBot \cpInfPlus1 \cpInfPlus2\end{center}
  \vspace*{-.5\baselineskip}
  \begin{center}\cpInfWith \cpInfNil \cpInfTop \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}\noindent

@@

%\section{Conclusions}
%\label{sec:conclusions}
%
%We introduced Hypersequent Classical Processes, a process calculus which addresses the mismatch between proof terms and \textpi-calculus terms which exists in Classical Processes~\citep{wadler2012}. We showed that \hcp enjoys preservation, progress and termination, and described its relation to \cp.
